Wavelength convergence of the $\exb$ staircase pattern in flux tube simulations of ion temperature gradient driven turbulence

### 2023-03-21 briefcommunicaton/briefcommunication.tex
#37: Binormal scan paragraph, graphic placement, threshold analysis

%% Include graphic for one column with specific place 
\newcommand{\includegraphicsOneCol}[3]{
	\begin{figure}[hbt!]
		\centering
		\includegraphics[width=0.9\linewidth]{#1}
		\caption{#2}
		\label{#3}
  	\end{figure}
  	\increasecounter{fig}{1}
}

@@

\newcommand{\wexb}{\omega_{\mathrm{\:E \times B}}}
\newcommand{\hatwexb}{\widehat{\omega}_{\mathrm{\:E \times B}}}
\newcommand{\exb}{\mathrm{\:E}\times\mathrm{B}}
\newcommand{\hatwexbkvec}{\widehat{\omega}_{\mathrm{\:E \times B},\mathbf{k}}}
%\newcommand{\NR}{N_\mathrm{R}}
%\newcommand{\NB}{N_\mathrm{B}}
\newcommand{\NR}{N_\psi}
\newcommand{\NB}{N_\zeta}
\newcommand{\rlt}{R/L_T}
%\newcommand{\rhoth}{\rho_\mathrm{th}}
\newcommand{\rhoth}{\rho}
\newcommand{\vth}{v_{\mathrm{th}}}
\newcommand{\nzf}{n_\mathrm{ZF}}
\newcommand{\kzf}{k_\mathrm{ZF}}

@@

The increase of the box size, relative to the standard box size $(L_\psi,~L_\zeta) = (76.27,~89.76)\,\rhoth$, in the radial and binormal direction is always indicated by the real parameter $\NR$ and $\NB$, respectively.
Here, the nomenclature $\NR\times \NB$ is applied throughout this work to indicate the increase in box size.  
Note that the number of modes in the respective direction is always adapted accordingly to retain a spatial resolution compliant to the standard resolution [Tab.~\ref{tab:resolution}] and standard box size. \bigskip

The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, rath2016, peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \psi^2},
	\label{eq:shearingrate}
\end{equation}
\increasecounter{eq}{1}{1}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature, $e$ is the elementary charge) and $\psi$ is the radial coordinate that labels the flux surfaces normalized by $\rhoth$. Consistent with Ref.~\cite{peeters2016} the turbulence level is quantified by the turbulent heat conduction coefficient $\chi$, which is normalized by $\rhoth^2 \vth/R$ ($\vth = \sqrt{2 T/m}$ is the thermal velocity and $m$ is the mass). Furthermore, quantities $\rhoth$, $R$, $T$, $\vth$ and $m$ are referenced quantities from Ref(\textbf{?}).
The zonal potential is estimated from the electrostatic potential $\phi$ on the two-dimensional $\psi$-$\zeta$-plane at the low field side according to
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\zeta} \int_0^{L_\zeta} \mathrm{d}\zeta ~ \phi(\psi,\zeta,s=0),
\end{equation}
\increasecounter{eq}{1}{1}
where $\zeta$ labels the field lines and is an approximate binormal or toroidal coordinate, and $s$ parameterizes the length along the field lines and represents a poloidal coordinate which represents together with the radial coordinate $\psi$ the Hamada coordinates $(\psi,\zeta,s)$. \cite{rath2021}(\textbf{?})
The E$\times$B shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}.

In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the $\exb$ shearing rate is considered. 
It is defined by
\begin{equation}
	\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \psi),
	\label{eq:shearingrate_fourier}
\end{equation}
\increasecounter{eq}{1}{1}
where $\hatwexb$ is the complex Fourier coefficient and \linebreak $\kzf = 2\pi \nzf/L_\psi$
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $\nzf$ ranging in $-(N_\psi -1)/2 \leq \nzf \leq (N_\psi -1)/2 $.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the definitions above, the shear carried by the zonal flow mode with wave vector $\kzf$ is defined by $|\hatwexb|_{\kzf} = 2 |\hatwexb|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the $\exb$ staircase pattern, also referred to as the \textit{basic mode} of the pattern in this work,exhibits the maximum amplitude in the spectrum $|\hatwexb|_{\kzf}$. \bigskip

@@

\includegraphicsTwoCol{{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1-1.5-2-2.5-3-5_Ns16_Nvpar48_Nmu9_wexb_comparison}.pdf}{
	Comparison of shearing rate $\wexb$ for each box sizes scan averaged over time interval and the growth rate $\pm \gamma$ of the most unstable linear ITG driven Eigenmode. The staircase structures got shifted till alignment for better visibility.\\
	\begin{tabular}{l l l l l l l l l}
		\textbf{(a) radial:}    & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 1}$   & $\in [15000, 18000]$, & $t_{3\times 1}$ & $\in [43000, 45000]$, & $t_{4\times 1}$ & $\in [26000, 28000]$ \\
		\textbf{(b) isotropic:} & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 2}$   & $\in [2000, 3000]$,   & $t_{3\times 3}$ & $\in [2000, 3000]$    &                 &                      \\
		\textbf{(c) binormal:}  & $t_{3\times 1.5}$ & $\in [11000, 12000]$, & $t_{3\times 2.5}$ & $\in [2000, 3000]$,   & $t_{3\times 5}$ & $\in [2000, 4000]$    &                 &                      \\
	\end{tabular}
}{fig:wexb-stable-comparison}{}

@@

% what kind of analysis
In a third test the binormal box size is varied with the radial box size fixed to $\NR = 3$.
This test covers the realizations $\NR \times \NB \in [3\times1.5,~3\times2.5,~3\times3,~3\times5]$. As in the isotropic scan the turbulence subdued and a fully developed staircase pattern formats always after $\sim 1000\,R/\vth$ although the turbulence of realization $3\times 1.5$ does suppress after $\sim 11000\,R/\vth$ [Fig. \ref{fig:eflux-3x1.5-2.5-3-5-comparison}]. The convergence of staircase pattern can be seen in Fig \ref{fig:wexb-stable-comparison}\,(c) and the test confirms again the typical mesoscale distinct from $\NR > 1$. Fig. \ref{fig:wexb-stable-comparison}\,(c) also confirms that indeed a competition between patterns with two wavelengths $\lambda \in [57.20,~ 76.27]\,\rhoth$ causing the different results for $3 \times 1$ and $3\times 3$. The zonal flow mode number varies between $\nzf = [3,4]$ which can be seen in Fig. \ref{fig:wexb-stable-comparison}\,(c) for the realization $3\times 2.5$ where the staircase structure has an (PATTERN). The Fourier analysis yields in this case no definitely basic mode rather than two dominate modes with $\nzf = 3, 4$ which split the maximum amplitude $|\hatwexb|_{\kzf}$ between them.

\includegraphicsOneCol{{Comparison/Boxsize/S6_rlt6.0_boxsize1x1-2x2-3x3_Ns16_Nvpar48_Nmu9_eflux_comparison}.pdf}{
	Time traces of the heat conduction coefficient $\chi$ for $\rlt = 6.0$ for radial and binormal increased box sizes
}{fig:eflux-1x1-2x2-3x3-comparison}

@@

% what kind of test
In the final test the inverse background temperature gradient length $\rlt$ is varied at fixed $3\times3$ box size.
Since suppression of turbulence usually occurs at later times when approaching the finite heat flux threshold from below \cite{peeters2016}, the analysis aims to lengthen the phase during which the zonal flow varies in time due to turbulent Reynolds stresses.
This scan covers realizations with $\rlt \in [6.0,~6.2,~6.4]$.
% what is the finite heat flux threshold
In the case of $\rlt = ???$ turbulence suppression is observed for $t > ???$, while stationary turbulence during the entire simulation time trace of $???~R/\vth$ is found for $\rlt = ???$.
The finite heat flux threshold, hence, is $\rlt|_\mathrm{finite} = ??? \pm ???$.
% Which stationay wavelength occurs????
Although the initial quasi-stationary turbulence in the former case is significantly larger compared to the $\rlt = 6.0$ realization discussed in the second test, a stationary pattern with basic zonal flow mode $\nzf = ???$ establishes. 
% Does it confirm our previous tests
Again, the $\nzf = 1$ (box scale) zonal flow mode does not grow secularly during the entire turbulent phase.
Also, this test confirms the statistical soundness of the converged pattern wavelength of $\sim ???$.\bigskip

>>>>

%% Include graphic for one column with specific place 
\newcommand{\includegraphicsOneCol}[3]{
	\begin{center}
		\centering
		\captionsetup{type=figure}
		\includegraphics[width=0.9\linewidth]{#1}
		\captionof{figure}{#2}
		\label{#3}
	\end{center}
  	\increasecounter{fig}{1}
}

%\newcommand{\includegraphicsOneCol}[3]{
%	\begin{figure}[hbt!]
%		\centering
%		\includegraphics[width=0.9\linewidth]{#1}
%		\caption{#2}
%		\label{#3}
%  	\end{figure}
%  	\increasecounter{fig}{1}
%}

@@

\newcommand{\wexb}{\omega_{\mathrm{\:E \times B}}}
\newcommand{\hatwexb}{\widehat{\omega}_{\mathrm{\:E \times B}}}
\newcommand{\exb}{\mathrm{\:E}\times\mathrm{B}}
\newcommand{\hatwexbkvec}{\widehat{\omega}_{\mathrm{\:E \times B},\mathbf{k}}}
%\newcommand{\NR}{N_\mathrm{R}}
%\newcommand{\NB}{N_\mathrm{B}}
\newcommand{\NR}{N_\mathrm{R}}
\newcommand{\NB}{N_\mathrm{B}}
\newcommand{\rlt}{R/L_T}
%\newcommand{\rhoth}{\rho_\mathrm{th}}
\newcommand{\rhoth}{\rho}
\newcommand{\vth}{v_{\mathrm{th}}}
\newcommand{\nzf}{n_\mathrm{ZF}}
\newcommand{\kzf}{k_\mathrm{ZF}}
\newcommand{\xcoord}{x}
\newcommand{\ycoord}{y}

@@

The increase of the box size, relative to the standard box size $(L_\xcoord,~L_\ycoord) = (76.27,~89.76)\,\rhoth$, in the radial and binormal direction is always indicated by the real parameter $\NR$ and $\NB$, respectively.
Here, the nomenclature $\NR\times \NB$ is applied throughout this work to indicate the increase in box size.  
Note that the number of modes in the respective direction is always adapted accordingly to retain a spatial resolution compliant to the standard resolution [Tab.~\ref{tab:resolution}] and standard box size. \bigskip

The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, rath2016, peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2},
	\label{eq:shearingrate}
\end{equation}
\increasecounter{eq}{1}{1}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature, $e$ is the elementary charge) and $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by $\rhoth$. Consistent with Ref.~\cite{peeters2016} the turbulence level is quantified by the turbulent heat conduction coefficient $\chi$, which is normalized by $\rhoth^2 \vth/R$ ($\vth = \sqrt{2 T/m}$ is the thermal velocity and $m$ is the mass). Furthermore, quantities $\rhoth$, $R$, $T$, $\vth$ and $m$ are referenced quantities from Ref(\textbf{?}).
The zonal potential is estimated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0),
\end{equation}
\increasecounter{eq}{1}{1}
where $\ycoord$ labels the field lines and is an approximate binormal or toroidal coordinate, and $s$ parameterizes the length along the field lines and represents a poloidal coordinate which represents together with the radial coordinate $\xcoord$ the Hamada coordinates $(\xcoord,\ycoord,s)$. \cite{rath2021}(\textbf{?})
The E$\times$B shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}.

In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the $\exb$ shearing rate is considered. 
It is defined by
\begin{equation}
	\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \xcoord),
	\label{eq:shearingrate_fourier}
\end{equation}
\increasecounter{eq}{1}{1}
where $\hatwexb$ is the complex Fourier coefficient and \linebreak $\kzf = 2\pi \nzf/L_\xcoord$
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $\nzf$ ranging in $-(N_\xcoord -1)/2 \leq \nzf \leq (N_\xcoord -1)/2 $.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the definitions above, the shear carried by the zonal flow mode with wave vector $\kzf$ is defined by $|\hatwexb|_{\kzf} = 2 |\hatwexb|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the $\exb$ staircase pattern, also referred to as the \textit{basic mode} of the pattern in this work,exhibits the maximum amplitude in the spectrum $|\hatwexb|_{\kzf}$. \bigskip

@@

\includegraphicsTwoCol{{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1-1.5-2-2.5-3-5_Ns16_Nvpar48_Nmu9_wexb_comparison}.pdf}{
	Comparison of shearing rate $\wexb$ for each box sizes scan averaged over time interval and the growth rate $\pm \gamma$ of the most unstable linear ITG driven Eigenmode. The staircase structures got shifted till alignment for better visibility.\\
	\begin{tabular}{l l l l l l l l l}
		\textbf{(a) radial:}    & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 1}$   & $\in [15000, 18000]$, & $t_{3\times 1}$ & $\in [43000, 45000]$, & $t_{4\times 1}$ & $\in [26000, 28000]$ \\
		\textbf{(b) isotropic:} & $t_{1\times 1}$   & $\in [2000, 5000]$,   & $t_{2\times 2}$   & $\in [2000, 3000]$,   & $t_{3\times 3}$ & $\in [2000, 3000]$    &                 &                      \\
		\textbf{(c) binormal:}  & $t_{3\times 1.5}$ & $\in [2000, 3000]$, & $t_{3\times 2.5}$ & $\in [2000, 3000]$,   & $t_{3\times 5}$ & $\in [1000, 3000]$    &                 &                      \\
	\end{tabular}
}{fig:wexb-stable-comparison}{}

\includegraphicsOneCol{{Comparison/Boxsize/S6_rlt6.0_boxsize1x1-2x2-3x3_Ns16_Nvpar48_Nmu9_eflux_comparison}.pdf}{
	Time traces of the heat conduction coefficient $\chi$ for $\rlt = 6.0$ for isotropic increased box sizes
}{fig:eflux-1x1-2x2-3x3-comparison}

@@

% what kind of analysis
In a third test the binormal box size is varied with the radial box size fixed to $\NR = 3$.
This test covers the realizations $\NR \times \NB \in [3\times1.5,~3\times2.5,~3\times3,~3\times5]$. As in the isotropic scan the turbulence subdued and a fully developed staircase pattern formats always after $\sim 2000\,R/\vth$ [Fig.~\ref{fig:eflux-3x1.5-2.5-3-5-comparison}]. The convergence of staircase pattern can be seen in Fig.~\ref{fig:wexb-stable-comparison}\,(c) and the test confirms again the typical mesoscale distinct from $\NR > 1$. Fig.~\ref{fig:wexb-stable-comparison}\,(c) also confirms that indeed a competition between patterns with two wavelengths $\lambda \in [57.20,~ 76.27]\,\rhoth$ causing the different results for $3 \times 1$ and $3\times 3$. The zonal flow mode number varies between $\nzf = [3,4]$ which can be seen in Fig.~\ref{fig:wexb-stable-comparison}\,(c). For the realization $3\times 2.5$ the staircase structure has an pattern between $3$ and $4$ repetitions which get represented in the second repetition with no $+ \gamma$ interval. Instead the pattern returns immediately after reaching $+ \gamma$ to the $- \gamma$ interval of the third repetition. The Fourier analysis yields in this case no definitely basic mode rather two dominate modes with $\nzf = 3, 4$ with a fraction of the maximum amplitude $|\hatwexb|_{\kzf}$ each.

@@

% what kind of test
In the final test the inverse background temperature gradient length $\rlt$ is varied at fixed $3\times3$ box size.
Since suppression of turbulence usually occurs at later times when approaching the finite heat flux threshold from below \cite{peeters2016}, the analysis aims to lengthen the phase during which the zonal flow varies in time due to turbulent Reynolds stresses.
This scan covers realizations with $\rlt \in [6.0,~6.2,~6.4]$.
% what is the finite heat flux threshold
In the case of $\rlt = 6.2$ turbulence suppression is observed for $t > 12000\,R/\vth$, while stationary turbulence during the entire simulation time trace of $12000\,R/\vth$ is found for $\rlt = 6.4$.
The finite heat flux threshold, hence, is $\rlt|_\mathrm{finite} = 6.4 \pm 0.1$.
% Which stationay wavelength occurs????
Although the initial quasi-stationary turbulence in the former case is significantly larger compared to the $\rlt = 6.0$ realization discussed in the second test, a stationary pattern with basic zonal flow mode $\nzf = 3$ establishes. 
% Does it confirm our previous tests
Again, the $\nzf = 1$ (box scale) zonal flow mode does not grow secularly during the entire turbulent phase.
Also, this test confirms the statistical soundness of the converged pattern wavelength of $\sim 57.20 - 76.27\,\rhoth$.\bigskip
